---
title: Real Time Earthquake Data
author: Péter Jeszenszky
date: September 17, 2022
output:
  github_document:
    number_sections: true
lang: en
---

# Data Source

The [U.S. Geological Survey](https://www.usgs.gov/) publishes real time earthquake data on the website of the [USGS Earthquake Hazards Program](https://earthquake.usgs.gov/). The data is available [here](https://earthquake.usgs.gov/earthquakes/feed/v1.0/csv.php) in CSV format.

# Reading Data

Let's read the CSV file from [here](https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/4.5_week.csv) that contains data about all M4.5+ earthquakes of the past 7 days:
```{r}
library(readr)
earthquakes <- read_csv("https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/4.5_week.csv",
  col_types = list(nst = col_integer())
)
```

Let's examine the structure of the dataset:
```{r}
str(earthquakes)
```

The dataset contains a total number of `r nrow(earthquakes)` rows:
```{r}
earthquakes
```

# Visualization

Let's visualize the earthquakes on a map such that the color of an earthquake should show the magnitude:
```{r}
library(ggplot2)
library(ggspatial)
ggplot(earthquakes, aes(longitude, latitude, color = mag)) +
  annotation_map_tile() +
  geom_spatial_point()
```

Histogram of the depth of the earthquakes:
```{r}
ggplot(earthquakes, aes(x = depth)) +
  geom_histogram(fill = "coral") +
  labs(
    title = "M4.5+ earthquakes of the past 7 days by depth",
    subtitle = "Source: USGS Earthquake Hazards Program"
  )
```

